### 2019-03-04 report/final-report.tex
Checkpoint

  \large \textbf{Synthesizing Data Wranglers\\\vspace{1cm} FA8750-17-2-0028}

@@

  \normalsize
  %\begin{tabular}{rl}
  Kathleen Fisher (Tufts University) \\
  Benjamin Pierce  (University of Pennsylvania) \\
  David Walker (Princeton University) \\
  Steve Zdancewic (University of Pennsylvania)
 %\end{tabular}
  \vspace{1in}

@@

Maintaining software systems that manage evolving representations of
data is tedious and error prone.  In this study, we have developed
algorithms for automatically synthesizing \emph{lenses}, which are
software adaptors that convert between two different representations
of the same information.  Lenses are a special form of program that
guarantee so-called \emph{round-tripping properties} that ensure the
transformations are well-behaved.  Our synthesis algorithms take as
input (i) the \emph{type} of each source as a regular expression, as
well as (ii) a small collection of representative \emph{examples} of
the desired translation.  They produceas output \emph{bijective string
  lenses}.  Bijective string lenses are a special class of lenses that
only exist when the two data sources are related via a bijection.
Despite this limitation, bijections and bijective synthesis form a
useful foundation on top of which more general transformations between
data sets may be built.  We explore two of these extensions: quotient
lenses and symmetric lenses.  Quotient lenses apply when two formats
are ``morally'' but not technically bijective.  Examples include two
formats that differ in their use of white space or in the order of
sequences of elements.  Symmetric lenses allow one side or the other to have
information not present in the other.  When converting between the
formats, defaults take the place of missing information.

@@

We have written
papers~\cite{bijective-synthesis,quotient-synthesis,symmetric-synthesis}
describing and evaluating our algorithms. We produced open-source 
code~\cite{GitHub} that implements those algorithms.  In this report,
we summarize our key methods, results, and findings, focusing
on the core algorithms for bijective synthesis, quotient synthesis and
simple symmetric synthesis.

>>>>

@@

  \Large \textbf{Synthesizing Data Wranglers\\\vspace{1cm} FA8750-17-2-0028}

  \large
  %\begin{tabular}{rl}
  Kathleen Fisher (Tufts University) \\
  Benjamin Pierce  (University of Pennsylvania) \\
  David Walker (Princeton University) \\
  Steve Zdancewic (University of Pennsylvania)
 %\end{tabular}
  \vspace{1in}

@@

Maintaining software systems that manage evolving representations of
data is tedious and error prone.  In this study, we have developed
algorithms for automatically synthesizing \emph{lenses}, which are
software adaptors that convert between two different representations
of the same information~\cite{Focal2005}.  Lenses are a special form of
programs that 
guarantee so-called \emph{round-tripping properties} that ensure the
transformations are well-behaved.  Our synthesis algorithms take as
input (i) the \emph{type} of each source as a regular expression, and
(ii) a small collection of representative \emph{examples} of
the desired translation.  They produce as output different varieties of \emph{string
  lenses}.  Have investigated three specific varieties in detail: {\em
  bijective lenses}, {\em quotient
lenses} and {\em simple symmetric lenses}.  Bijective lenses address the
simplest case, when the two formats we want to convert between have exactly
the same information content, modulo rearrangement and reformatting.  
Quotient lenses apply when two
formats are ``morally'' but not technically bijective.  Examples include two
formats that differ in their use of white space or in the order of
sequences of elements.  Simple symmetric lenses allow one side or the other
to have 
information not present in the other; when converting between the
formats, defaults take the place of missing information.

@@

The products of this investigation include several published
papers~\cite{bijective-synthesis, quotient-synthesis, symmetric-synthesis}
describing and evaluating our algorithms, as well as open-source 
code~\cite{GitHub} that implements those algorithms.  In this report,
we summarize our key methods, results, and findings, focusing
on the core algorithms for bijective synthesis, quotient synthesis, and
simple symmetric synthesis.
